Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-04-11 emsoft-2019/Motiv.tex
two more comments
disturbance

% !TEX root = main.tex
\section{Overview}\label{sec:example}

@@

As a smaller example, consider the standard problem of designing a controller for an inverted pendulum depicted in~\autoref{fig:inverted_pendulum}~{(a)}.
The goal of the controller is to keep the pendulum at the vertical position while satisfying hard constraints on the state variables and control inputs.
A model of the system can be constructed using physical principles. 
After linearization and time discretization, the model is
	\begin{equation}
		\begin{bmatrix}
			 \theta_{k+1}\\
			\omega_{k+1}
		\end{bmatrix}=
		\begin{bmatrix}
			1 & T_s\\
			\frac{T_sg}{L}& (1-\frac{T_sb}{mL^2})		
		\end{bmatrix}
		\begin{bmatrix}
			\theta_k\\
			\omega_k
		\end{bmatrix}+
		\begin{bmatrix}
			0\\
			\frac{T_s}{mL^2}
		\end{bmatrix}u_k + w_k
		\label{eq:pendul_ss}
	\end{equation}
where $\theta_k$, $\omega_k$, and $u_k$ denote respectively the angular position, angular speed, 
and the input torque at time $kT_s$ with $T_s$ being an appropriate sample time.
The disturbance $w_k\in\mathbb R^2$, bounded by $\|w_k\|\leq \Delta$, captures the modeling error due to 
linearization and discretization. 
The parameter $g=9.81 [m/s^2]$ is the gravitational acceleration, $m$ is the ball mass, $b$ is the rotational fraction coefficient, 
and $L$ is the length of the bar. 
Starting from an initial state $(\theta_0,\omega_0)$, the control goal is to converge to the equilibrium point 
$\theta=0, \omega=0$.
Additionally, we require the state constraints $\theta_k\in[-\pi,\pi]$ and $\omega_k\in[-\pi/8,\pi/8]$ to hold at all time instances.

Our overall goal is to 
(i) design a robust MPC controller that achieves the performance objectives including hard constraints in spite of a bounded
disturbance $\Delta$ modeling the implementation error due to finite precision;
and at the same time
\eva{what about the modelling error that $\Delta$ bounded in the paragraph before?}
(ii) to minimize the total number of bits that are required for storing the explicit controller 
while ensuring the error $e\leq \Delta$ is satisfied by the implementation.

@@

Figure \ref{fig:overview} gives a high-level overview of our proposed setup. 
We start by defining an initial bound $\Delta_0$. 
For our example, we choose $\Delta_0=0$. 
We use MATLAB to find a robust explicit MPC with the bound $\Delta_0$ on the disturbance.
Next, we ask Daisy
to provide us with a mixed-precision scheme such that the implementation satisfies the error bound $\Delta$, that is,
at all points in the control domain, the difference between the (infinite precision) controller function
and its implementation in the mixed-precision scheme is at most $\Delta$.

Unfortunately, $\Delta_0=0$ is not realizable,  as the error resulting from a fixed-point 
implementation is always greater than zero. 
Therefore, we increase $\Delta$ by $0.05$ and try again. 
Solving the robust explicit MPC problem for the new disturbance bound gives the new controller,
shown in Figure~\ref{fig:inverted_pendulum}(b).
The controller has \RM{XX} regions.
\RM{give the control function. I think it will be good to explicitly write the polyhedra and the affine maps}

Next, we ask Daisy to provide us with mixed-precision implementation for all the parameters and to compute a static bound on the
implementation error. 
The mixed-precision scheme given by Daisy results in an implementation with $6084$ bits in total 
for the corresponding robust explicit MPC.
Daisy also computes a static error bound of \RM{what?}
Since this error bound is less than our robustness margin $\Delta = 0.05$, we terminate the loop
and return this implementation.
The mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta$,
and about $XXX\%$ \RM{what} less memory than an implementation that uniformly uses
32 bits.

>>>>

% !TEX root = main.tex
\section{Overview}
\label{sec:example}

@@

As a smaller example, consider the standard problem of designing a controller for an inverted pendulum depicted in~\autoref{fig:inverted_pendulum}~{(a)}.
The goal of the controller is to keep the pendulum at the vertical position while satisfying hard constraints on the state variables and control inputs.
A model of the system can be constructed using physical principles. 
After linearization and time discretization, the model is
	\begin{equation}
		\begin{bmatrix}
			 \theta_{k+1}\\
			\omega_{k+1}
		\end{bmatrix}=
		\begin{bmatrix}
			1 & T_s\\
			\frac{T_sg}{L}& (1-\frac{T_sb}{mL^2})		
		\end{bmatrix}
		\begin{bmatrix}
			\theta_k\\
			\omega_k
		\end{bmatrix}+
		\begin{bmatrix}
			0\\
			\frac{T_s}{mL^2}
		\end{bmatrix}u_k + w_k
		\label{eq:pendul_ss}
	\end{equation}
where $\theta_k$, $\omega_k$, and $u_k$ denote respectively the angular position, angular speed, 
and the input torque at time $kT_s$ with $T_s$ being an appropriate sample time.
The disturbance $w_k\in\mathbb R^2$, \Sadegh{bounded with a polyhedral set $w_k\in \mathcal W$, captures the modeling error due to 
linearization and discretization}. 
The parameter $g=9.81 [m/s^2]$ is the gravitational acceleration, $m$ is the ball mass, $b$ is the rotational fraction coefficient, 
and $L$ is the length of the bar. 
Starting from an initial state $(\theta_0,\omega_0)$, the control goal is to converge to the equilibrium point 
$\theta=0, \omega=0$.
Additionally, we require the state constraints $\theta_k\in[-\pi,\pi]$ and $\omega_k\in[-\pi/8,\pi/8]$ to hold at all time instances.

Our overall goal is to 
(i) design a robust MPC controller that achieves the performance objectives including hard constraints in spite of \Sadegh{an additional bounded
disturbance $\Delta$ modeling the implementation error (both errors due to choosing a wrong region and the finite-precision computation of the control action);}
and at the same time
%\eva{what about the modelling error that $\Delta$ bounded in the paragraph before?}
(ii) to minimize the total number of bits that are required for storing the explicit controller 
while ensuring the implementation error $e\leq \Delta$ is satisfied by the implementation.

@@

Figure \ref{fig:overview} gives a high-level overview of our proposed setup.
\Sadegh{
We start by selecting an initial bound $\Delta_0$ on the implementation error and enlarge $\mathcal W$ with $\Delta_0$ as $\mathcal W_{\Delta_0}$, where
\begin{equation}
\mathcal W_{\Delta} := \{w_1+w_2 \,|\, w_1\in \mathcal W, \|w_2\|\le \Delta\}
\quad\text{ for all } \Delta\ge 0.
\end{equation}
 }

 \Sadegh{
For our example, we choose $\Delta_0=0$ and $\mathcal W = \{0\}$. 
We use MATLAB to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$.
Next, we ask Daisy
to provide a mixed-precision scheme such that the associated implementation error is at most $\Delta_0$, that is,
at all points in the control domain, the difference between the infinite-precision controller function
and its mixed-precision implementation is at most $\Delta_0$.
}

\Sadegh{
Unfortunately, $\Delta_0=0$ is not realizable, as the implementation error is always greater than zero either due to selecting the wrong region in the affine controller function or due to the finite-precision computation of the control action.
%\eva{Is this the error due to mixed-precision or due to wrong regions?}
Therefore, we define $\Delta_1 := \Delta_0 + \delta$ with $\delta$ being a fixed threshold (we chose $\delta = 0.05$) and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller,
shown in Figure~\ref{fig:inverted_pendulum}(b).
The controller has \RM{XX} regions.
\RM{give the control function. I think it will be good to explicitly write the polyhedra and the affine maps} 
}

Next, we ask Daisy to provide us with mixed-precision implementation for all the parameters and to compute a static bound on the
implementation error. 
The mixed-precision scheme given by Daisy results in an implementation with $6084$ bits in total 
for the corresponding robust explicit MPC.
Daisy also computes a static error bound of \RM{what?}
Since this error bound is less than our robustness margin $\Delta_1 = 0.05$, we terminate the loop
and return this implementation.
The mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $XXX\%$ \RM{what} less memory than an implementation that uniformly uses
32 bits.
